Map/Reduce for the common man

### 2010-02-16 doc/slides.tex
*** empty log message ***

@@

\begin{frame}\frametitle{}
  common part between cores and network

  output/input\_value on core

  architecture-independent protocol for network

  same program for master and workers
\end{frame}

>>>>

  \begin{ocaml}
master_loop : 
  create_job:('worker -> 'task -> unit) -> 
  wait:(unit -> 'worker * 'task list) -> 
  'worker list -> 'task list -> unit
  \end{ocaml}
\end{frame}

\begin{frame}\frametitle{Cores}
  \begin{itemize}\item 
    \texttt{create\_job} is achieved by \texttt{Unix.fork}, using a
    local file to store the result (Unix pipes have size limitations)

  \item 
    \texttt{wait} is achieved using \texttt{Unix.wait}
  \end{itemize}

  Ocaml's marshalling is safely exploited to read/write values of any type

  scheduling tasks to various cores is left to the OS
\end{frame}

\begin{frame}\frametitle{Network}
  master and workers are the same program but distinguish using an
  environment variable (\texttt{WORKER})

  \begin{itemize}
  \item a worker starts a server, waiting for tasks
  \item the master connects to workers, sending tasks
  \end{itemize}

  simple, string-based protocol:
  \begin{center}
    \begin{tabular}{ccc}
      \framebox{master} &  & \framebox{worker} \\
      &     $\stackrel{\displaystyle connect}{\longrightarrow}$ &  \\
      &     $\stackrel{\displaystyle\mathtt{Assign}~3~\mathtt{"f"}~\mathtt{"7"}}
                      {\longrightarrow}$ &  \\
      &     $\stackrel{\displaystyle \mathtt{Started}~3}
                      {\longleftarrow}$ &  \\
      &     $\stackrel{\displaystyle 
                       \only<1>{\mathtt{Completed}~3~\mathtt{"42"}}
                       \only<2>{\emph{\mathtt{Aborted}~3~}}
                       \only<3>{\emph{\mathtt{Kill}~3}}}
                      {\only<1-2>{\longleftarrow}
                       \only<3>{\longrightarrow}}$ &  \\
    \end{tabular}
  \end{center}
\end{frame}

\begin{frame}\frametitle{Network: Polymorphic Map/Reduce}
  if master and workers are \emph{exactly} the same executable, then
  marshalling can still be used to pass values $\Rightarrow$
  polymorphic map/reduce

  if master and workers are compiled from the \emph{same source}, then
  we are limited to strings $\Rightarrow$
  monomorphic map/reduce

  \footnotesize
  \begin{ocaml}
module Str : sig

  val map : f:(string -> string) -> string list -> string list

  val map_local_reduce :
    map:(string -> string) -> 
    reduce:(string -> string -> string) -> 
    string -> string list -> string
  ...
  \end{ocaml}
\end{frame}

\begin{frame}\frametitle{Master}
  whatever the implementation is (cores or network), we now 
  provide the same generic master loop

  \begin{ocaml}
    master : (task -> task_result) -> 
             (task -> task_result -> task list) ->
             task list -> 
             unit
  \end{ocaml}

  (the notion of workers is no more explicit)
\end{frame}

\begin{frame}\frametitle{Implementing \texttt{map}}
  \begin{ocaml}
val map : f:('a -> 'b) -> 'a list -> 'b list
  \end{ocaml}

  \begin{enumerate}
  \item initially, the $i$-th element $x_i$ of the list is the task $(i,x_i)$
  \item when task $(i,x_i)$ completes, store the result $f(x_i)$ in a table
  \item when done, rebuild the list from the table contents
  \end{enumerate}
\end{frame}

\begin{frame}\frametitle{Implementing \texttt{map\_local\_reduce}}
  \begin{ocaml}
 val map_local_reduce :
  map:('a -> 'b) -> reduce:('c -> 'b -> 'c) -> 
  'c -> 'a list -> 'c
  \end{ocaml}

  \begin{enumerate}
  \item initially, each element $x_i$ of the list is a task
  \item when task $x_i$ completes, the result $f(x_i)$ is
    \emph{locally} combined with the current accumulator (a local reference)
  \item when done, return the accumulator
  \end{enumerate}
\end{frame}

\begin{frame}\frametitle{Implementing \texttt{map\_remote\_reduce}}
  \begin{ocaml}
val map_remote_reduce :
  map:('a -> 'b) -> reduce:('c -> 'b -> 'c) -> 
  'c -> 'a list -> 'c
  \end{ocaml}

  \begin{enumerate}
  \item initially, each element $x_i$ in the list is a task $Map(x_i)$
  \item when a task is completed with result $r$,
    \begin{itemize}
    \item either it is $Map(x_i)$,
      \begin{itemize}
      \item if accumulator is $\texttt{Some}~v$, generate the task
        $Reduce(v,r)$ and set accumulator to \texttt{None}
      \item if accumulator is \texttt{None}, push $r$ into $\mathit{stack}_r$
      \end{itemize}
    \item or it is $Reduce(r',v')$
      \begin{itemize}
      \item if $\mathit{stack}_r$ is empty, the accumulator is now
        $\texttt{Some}~r$ 
      \item otherwise, pop $v$ from $\mathit{stack}_r$, and generate
        $Reduce(r,v)$ 
      \end{itemize}
    \end{itemize}
  \item when done, return the accumulator
  \end{enumerate}
\end{frame}

\begin{frame}\frametitle{Implementing \texttt{map\_reduce\_ac}}
  \begin{ocaml}
val map_reduce_ac :
  map:('a -> 'b) -> reduce:('b -> 'b -> 'b) -> 
  'b -> 'a list -> 'b
  \end{ocaml}

  \begin{enumerate}
  \item initially, each element $x_i$ in the list is a task $Map(x_i)$
  \item when a task is completed with result $r$,
    \begin{itemize}
    \item if the accumulator is \texttt{None}, store $r$ in the accumulator
    \item else, combine $r$ with the accumulator $v$ into a new task
      $Reduce(v, r)$ and set accumulator to \texttt{None}
    \end{itemize}
  \item when done, return the accumulator
  \end{enumerate}
\end{frame}

\begin{frame}\frametitle{Implementing \texttt{map\_reduce\_a}}
  \begin{ocaml}
val map_reduce_a :
  map:('a -> 'b) -> reduce:('b -> 'b -> 'b) -> 
  'b -> 'a list -> 'b
  \end{ocaml}

  \begin{enumerate}
  \item initially, each element $x_i$ in the list is a task $(i,i,Map(x_i))$
  \item when a task $(i,j)$ is completed with result $r$,
    \begin{itemize}
    \item if there is some result adjacent to $(i,j)$ available,
      merge them into a new reduce task
    \item else, store it
    \end{itemize}
  \item when done, return the only remaining result
  \end{enumerate}
\end{frame}

\begin{frame}\frametitle{Independent Master and Workers}
  the library actually provides tools for independent development of
  master \small

  \begin{ocaml}
module Master : sig
  ...
  val map_remote_reduce : string -> string list -> string
  ...
end
  \end{ocaml}

and workers\small

  \begin{ocaml}
module Worker : sig
  val register_computation : 
    string -> (string -> string) -> unit
  val register_computation2 : 
    string -> (string -> string -> string) -> unit
  val compute : 
    ?stop:bool -> ?port:int -> unit -> string
end
  \end{ocaml}
\end{frame}

\begin{frame}\frametitle{Future Work}

@@

\begin{frame}\frametitle{}
  \begin{center}
    \emph{thank you!}
  \end{center}
\end{frame}
